\newpage
\section{Local Search}
Solve problems approximately --- aims at a local optimum. 

Framework of Local Search:
\begin{itemize}\small
    \item Loacl 
    \subitem Define neighborhoods in the feasible set. 
    \subitem A local optimum is a best solution in a neighborhood. 
    \item Search 
    \subitem Start with a feasible solution and search a better one within the neighborhood. 
    \subitem A local optimum is achieved if no improvement is possible. 
\end{itemize}

Neighbor Relation:
\begin{itemize}\small
    \item  $S \sim S'$ : $S'$ is a neighboring solution of $S$ --- $S'$ can be obtained by a small modification of $S$.
    \item  $N(S)$: neighborhood of $S$ --- the set $\{ S': S \sim S' \}$.
\end{itemize}
\begin{lstlisting}[language={c++}]
SolutionType Gradient_descent(){
    Start from a feasible solution S in FS;
    MinCost=cost(S);
    while(1){
        S_=Search(N(S)); /* find the best S_ in N(S) */
        CurrentCost=cost(S_);
        if(CurrentCost<MinCost){
            MinCost=CurrentCost;    
            S=S_;
        }
        else  break;
    }
    return S;
}
\end{lstlisting}

\subsection{The Vertex Cover Problem}
Given an undirected graph $G = (V, E)$. Find a minimum subset $S$ of $V$ such that for each edge $(u, v)$ in $E$, either $u$ or $v$  is in $S$.

\begin{itemize}
    \item Feasible solution set $FS$ : all the vertex covers. 
    \item $cost(S)=|S|$
    \item $S\sim S'$: Each vertex cover $S$ has at most $|V|$ neighbors
    \item Search: Start from $S = V$; delete a node and check if $S'$ is a vertex cover with a smaller cost.
\end{itemize}

\subsection{Simulated Annealing}
The material is cooled very gradually from a high temperature, allowing it enough time to reach equilibrium at a succession of intermediate lower temperatures.

\subsection{Hopfield Neural Networks}
Graph $G = (V, E)$ with integer edge weights $w$ (positive or negative).
\begin{itemize}
    \item If $w_e < 0$, where $e = (u, v)$, then $u$ and $v$ want to have the same state($\pm 1$). 
    \item if $w_e > 0$ then $u$ and $v$ want different states.
\end{itemize}
The absolute value $|w_e|$ indicates the strength of this requirement.

Find a sufficiently good configuration $S$ of the network --- an assignment of the state $s_u$ to each node $u$. 

\begin{definition}
    In a configuration $S$, edge $e = (u, v)$ is good if $w_e s_u s_v < 0$ ($w_e < 0$ iff $s_u = s_v$ ); otherwise, it is bad.
\end{definition}
\begin{definition}
    In a configuration $S$, a node $u$ is satisfied if the weight of incident good edges $\ge$ weight of incident bad edges.
    \begin{align*}
        \sum_{u,v:e=(u,v)\in E}w_es_us_v\le 0
    \end{align*}
\end{definition}
\begin{definition}
    A configuration is stable if all nodes are satisfied. 
\end{definition}

\subsubsection{State-flipping Algorithm}
\begin{lstlisting}[language={c++}]
ConfigType State_flipping(){
    Start from an arbitrary configuration S;
    while(!IsStable(S)){
        u = GetUnsatisfied(S);
        su = -su;
    }
    return S;
}
\end{lstlisting}
\begin{claim}
    The state-flipping algorithm terminates at a stable configuration after at most $\displaystyle W = \sum_e|w_e|$ iterations.
\end{claim}

Related to Local Search:
\begin{itemize}
    \item Problem:  To maximize $\Phi$.
    \item Feasible solution set FS: configurations
    \item $S\sim S'$: $S'$ can be obtained from $S$ by flipping a single state
\end{itemize}
\begin{claim}
    Any local maximum in the state-flipping algorithm to maximize $\Phi$ is a stable configuration.
\end{claim}

\subsection{The Maximum Cut Problem}
Given an undirected graph $G = (V, E)$ with positive integer edge weights $w_e$, find a node partition $(A, B)$ such that the total weight of edges crossing the cut is maximized.
\begin{align*}
    w(A,B):=\sum_{u\in A, v\in B}w_{uv}
\end{align*}

Related to Local Search:
\begin{itemize}
    \item Problem:  To maximize $\displaystyle \Phi(S)=\sum_{e\text{ is good}}|w_e|$.
    \item Feasible solution set FS: any partition $(A, B)$ 
    \item $S\sim S'$: $S'$ can be obtained from $S$ by moving one node from $A$ to $B$, or one from $B$ to $A$.
\end{itemize}
A special case of Hopfield Neural Network with $w_e$ all being positive. 

\begin{claim}
    Let $(A, B)$ be a local optimal partition and let $(A^*, B^*)$ be a global optimal partition.  Then $w(A, B) \ge \frac{1}{2} w(A^*, B^*)$.
\end{claim}
Unless P = NP, no $17/16$(1.0625) approximation algorithm for MAX-CUT.

\subsubsection{Big-improvement-flip}
Only choose a node which, when flipped, increases the cut value by at least
\begin{align*}
    \frac{2\varepsilon}{|V|}w(A,B)
\end{align*}
\begin{claim}
    Upon termination, the big-improvement-flip algorithm returns a cut $(A, B)$ so that
    \begin{align*}
        (2+\varepsilon)w(A, B)\ge w(A^*, B^*)
    \end{align*}
\end{claim}

Since $\forall x\ge 1$, we have $\left( 1+\frac{1}{x} \right)^x\ge 2$, so the objective function doubles at least every $\frac{n}{\varepsilon}$  flips.

\begin{claim}
    The big-improvement-flip algorithm terminates after at most $O\left( \frac{n}{\varepsilon}\log W \right)$ flips.
\end{claim}

\subsubsection{K-L heuristic}
A better local, k-flip. 
\begin{enumerate}
    \item [Step 1] make 1-flip as good as we can - $O(n)$ $\Rightarrow$ $(A_1, B_1)$ and $v_1$
    \item [Step k] make 1-flip of an unmarked node as good as we can - $O(n-k+1)$ $\Rightarrow$ $(A_k, B_k)$ and $v_1, \dots, v_k$
    \item [Step n] $(A_n, B_n)=(B, A)$
\end{enumerate}
Neighborhood of $(A, B) = \{ (A_1, B_1), \dots, (A_{n-1}, B_{n-1}) \}$ - $O(n^2)$
