\newpage
\section{Inverted Files Index}

\textbf{Solution:}
\begin{enumerate}
    \item scan string
    \item Term-Document Incidence Matrix. e.g. Document sets  
\end{enumerate}

\subsection{Compact Version - Inverted File Index}
\begin{definition}
    \hl{Index} is a mechanism for locating a given term in a text. 
\end{definition}
\begin{definition}
    \hl{Inverted file} contains a list of pointers (e.g. the nuber of the page) to all occurrences of that term in the text. 
\end{definition}

Assume: bag of words (词顺序没那么重要)

\subsubsection{Index Genertor}
\begin{algorithm}[H]
    \caption{Index Genertor}
    \begin{algorithmic}
        \ForAll{document D}
            \ForAll{term T in D}
                \If{T doesn't exist in D}
                    \State insert T in D
                \EndIf
                \State Get T's posting list. 
                \State Insert a node to T's posting list. 
            \EndFor
        \EndFor
    \end{algorithmic}
\end{algorithm}

\subsubsection{While reading a term}
\begin{enumerate}
    \item \textbf{Word Stemming:} Process a word so that only its stem or root form is left. 
    \item \textbf{Stop Words:} Some words are so common that almost every document contains them, such as ``a'' ``the'' ``it''.  It is useless to index them.  They are called stop words.  We can eliminate them from the original documents.
\end{enumerate}

\subsubsection{While accessing a term}
\textbf{Solution:}
\begin{enumerate}
    \item Search trees
    \item Hashing
\end{enumerate}

\subsubsection{While not having enough memory}
Memory in blocks (分块存储), and merge them. 

Distributed indexing --- Each node contains index of a subset of collection. 

\textbf{Solution:}
\begin{enumerate}
    \item Term-partitioned index
    \item Document-partitioned indedx (robust)
\end{enumerate}

\subsubsection{Dynamic indexing}

\begin{itemize}
    \item Docs come in over time
    \begin{itemize}
        \item postings update for terms already in dictionary. 
        \item new terms added to dictionary. 
    \end{itemize}
    \item Docs get deleted
\end{itemize}

Using main index and auxiliary index. 

\subsubsection{Compression}
差分压缩index

\subsubsection{Thresholding}
Document: only retrieve the top $x$ documents where the documents are ranked by weight. 

\textbf{Problems:}
\begin{itemize}
    \item Not feasible for Boolean queries. 
    \item Can miss some relevant documents due to truncation. 
\end{itemize}

Query: Sort the query terms by their frequency in ascending order; search according to only some percentage of the original query terms. 

\subsection{Measures for a search engine}
\begin{itemize}
    \item How fast does it index
    \item How fast does it search
    \item Expressiveness of query language
\end{itemize}

\subsubsection{Relevance measurement}
\begin{enumerate}
    \item A benchmark document collection
    \item A benchmark suite of queries
    \item A binary assessment of either Relevant or Irrelevant for each query-doc pair   
\end{enumerate}
\begin{table}[H]
    \centering
    \begin{tabular}[c]{|r|c|c|}\hline
        & Relevant & Irrelevant\\ \hline
        Retrieved & $R_R$ & $I_R$\\ \hline
        Not Retrieved & $R_N$ & $I_N$ \\  \hline
    \end{tabular}
\end{table}
\begin{align*}
    \text{Precision }P&=\frac{R_R}{R_R+I_R}\\
    \text{Recall }R&=\frac{R_R}{R_R+R_N}
\end{align*}