\newpage
\section{Leftist Heaps and Skew Heaps}

\subsection{Leftist Heaps}
\begin{itemize}
    \item [\textbf{Target:}] Speed up merging in $O(N)$.
    \item [Heap:] Structure Property + Order Property
\end{itemize} 
Leftist Heap:
\begin{itemize}
    \item Order Property --- the same
    \item Structure Property --- binary tree, but unbalanced
\end{itemize}

\begin{definition}
    The \hl{null path length, NPl($X$) (距离)}, of any node $X$ is the length of the shortest path from $X$ to a node without two children. Define Npl(NULL)=-1. 
\end{definition}
Note: Npl($X$)=min\{Npl($C$)+1 for all $C$ as children of $X$\}

\begin{definition}
    The \hl{leftist heap property} is that for every node $X$ in the heap, the null path length of the \textcolor{light_green}{left} child is at least \hl{as large as} that of the \textcolor{light_green}{right} child. 
\end{definition}

\begin{theorem}
    A leftist tree with $r$ nodes on the right path must have at least $2^r-1$ nodes.
\end{theorem}

Note: The leftist tree of N nodes has a right path containing at most $\lfloor\log(N+1)\rfloor$ nodes. 

\begin{lstlisting}[language=c++,title={Declaration}]
struct TreeNode{
    ElementType Element;
    PriorityQueue Left;
    PriorityQueue Right;
    int Npl;
};
\end{lstlisting}

\subsubsection{Merge (recursive version)}
Have two leftist tree $H_1$ and $H_2$. 
\begin{enumerate}
    \item Merge($H_1$->Right, $H_2$), the top of $H_1$ is less than $H_2$
    \item Attach($H_2$, $H_1$->Right), attach $H_2$ to $H_1$->Right 
    \item Swap($H_1$->Right, $H_1$->Left)\\ if Nlp(Left) < Nlp(Right) 
\end{enumerate}

\subsubsection{Merge (iteration version)}
Have two leftist tree $H_1$ and $H_2$. 
\begin{enumerate}
    \item Sort the right paths without changing their left children. 
    \item Swap children if Nlp(Left) < Nlp(Right)
\end{enumerate}

$T_p=O(\log N)$

\subsubsection{Delete Min}
\begin{enumerate}
    \item Delete teh root
    \item Merge the two subtrees. 
\end{enumerate}

$T_p=O(\log N)$

\subsection{Skew Heaps}
\begin{itemize}\small
    \item [\textbf{Target}:] Any $M$ consecutive operations take at most $O(M\log N)$ time. 
    \item [Merge:] \hl{Always} swap the left and right children  except the \hl{largest} of all the nodes on the right paths doesn't have its children swapped. No Npl. 
\end{itemize}

Skew heaps have the advantage that no extra space is required to maintain path lengths and no tests are required to determine when to swap children.

It is an open problem to determine precisely the expected right path length of both leftist and skew heaps.

\subsubsection{Amortized Analysis for Skew Heaps}
Insert \& Delete are just Merge
\begin{itemize}
    \item $T_{amortized}=O(\log N)$ ?
    \item $D_i=$ the root of the resulting tree 
    \item $\Phi(D_i)=$ number of \hl{heavy} nodes
\end{itemize}
\begin{definition}
    A node $p$ is \hl{heavy} if the number of descendants of $p$'s right subtree is at least half of the number descendants of $p$, and \textcolor{light_green}{light} otherwise. The number of descendants of a node includes the node itself. 
\end{definition}

The only nodes whose heavy/light status can change are nodes that are initially on the right path.
\begin{align*}
    H_i=l_i+h_i\ (i=1,2),\ T_{worst}=l_1+h_1+l_2+h_2
\end{align*}

Before merge: $\Phi_i=h_1+h_2+h$

After merge: $\Phi_{i+1} \le l_1+l_2 +h$

\begin{align*}
    T_{amortized}=T_{worst}+\Phi_{i+1}-\Phi_i\le 2(l_1+l_2)
\end{align*}

$\because\ l=O(\log N)\ \therefore\ T_{amortized}=O(\log N)$