\newpage
\section{Week 1}

\subsection{框架}
划分问题的难度. 
\subsubsection{Automata Theory}
讨论问题前需要有清晰的数学定义. Definition of problems and computing models. 

finite automata ->pushdown automata -> Turing machines 

有个 Church-Turing Thesis, 断言 Turing machines 是最终极的模型. 

\subsubsection{Computability Theory}
可计算性问题. 

\subsubsection{Complexity Theory}
证明 hardness, 需要证明问题属于某个 complexity class (复杂类). 


\subsection{Problem Definition}

\subsubsection{Optimization problem}
e.g. Given $G:(V,E,w)$, what is the minimum spanning tree. 

\subsubsection{Search problem}
e.g. Given $G$ and an integer $k$, find a spanning tree whose weight is at most $k$ or tells such a tree not exist. 

\subsubsection{Decision problem}
e.g. Given $G$ and $k$, is there a spanning tree with weight at most $k$. 

\subsubsection{Counting problem}
e.g. Given $G$ and $k$, what is \# (the number of) spanning tree with weight at mots $k$?

Decision problem 最简单, 因为能解决任意其他三个问题, 就能解决 decision problem. 所有课程着眼于 decision problem, 接下来问题都是 decision problem. 

\subsection{抽象问题为数学形式语言}
Decision 结果为 yes-instance or no-instance. 

For computer 上述问题等价为: Given a string $w$, is $w \in $
\begin{align*}
    \{ \text{encode } (G, k): (G, k) \text{ is a yes-instance} \}
\end{align*}

问题由集合唯一决定, 称集合为 a language. 

\subsection{Language Definition}

\begin{definition}[Alphabet]
    A alphabet (字符集) is a \textbf{finite} set of symbols. 
\end{definition}

e.g. $\Sigma = \{0,1\}$, $\Sigma=\{  \}$ (空集). 


\begin{definition}[String]
    A string is a \textbf{finite} sequence of symbols from some alphabet. 
\end{definition}

e.g. $\Sigma=\{ 1,2,3 \},\ 1,3,23, 333$

Length $|w|$= \# symbols in $w$. 

Empty string $e$ with $|e|=0$. (特殊定义)

$\Sigma^i=$ the set of all string of length $i$ over $\Sigma$. 

e.g. $\Sigma=\{ 0, 1 \}$, $\Sigma^o=\{ e \}$, 
\begin{align*}
    \Sigma^*&=\bigcup_{i\ge 0}\Sigma^i\\
    \Sigma^+&= \bigcup_{i\ge 1} \Sigma^i
\end{align*}

\begin{definition}[Concatenation]
    Given $u=a_1\cdots a_n$, $v=b_1\cdots b_m$, $w=uv=a_1\cdots a_n b_1 \cdots b_m$
\end{definition}

\begin{definition}[exponentiation(幂)]
    $w^i=\underbrace{w\dots w}_{i \text{ times}}$
\end{definition} 

\begin{definition}[reversal]
    $w=a_1\dots a_n$, $w^R=a_n\dots a_1$. 
\end{definition}

\begin{definition}[Language]
    A language over $\Sigma$ is a subset $L \subseteq  \Sigma^*$. 
\end{definition}

e.g. $\Sigma=\{ 0,1 \}$, language is $\phi$, $\Sigma^*$, $\{ 0^n 1^n : n\ge 0 \}$. 

decision problem 与 language 一一对应. decision problem $\Leftrightarrow$ language. 
\begin{proof}
    decision problem $\Rightarrow$ \{ encoding of yes-instance \}. 

    Given a string $w$, is $w\in L$ $\Leftarrow$ language
\end{proof}

\subsection{Finite Automata}
有限状态机. state diagram

\begin{figure}[!htb]
    \centering
    \begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
        \node[state,initial]  (q_0)                      {$q_0$};
        \node[state]          (q_1) [above right=of q_0] {$q_1$};
        \node[state]          (q_2) [below right=of q_0] {$q_2$};
        \node[state,accepting](q_3) [below right=of q_1] {$q_3$};
    
        \path[->] (q_0) edge              node        {0} (q_1)
                        edge              node [swap] {1} (q_2)
                  (q_1) edge              node        {1} (q_3)
                        edge [loop left]  node        {0} ()
                  (q_2) edge              node [swap] {0} (q_3)
                        edge [loop above] node        {1} ();
    \end{tikzpicture}
    \caption{example state diagram}
\end{figure}

initial state is unique, Finial state may be several. 

\begin{definition}
    A finite automata $M=(K, \Sigma, \delta, s, F)$. 
    \begin{itemize}
        \item $\Sigma$: input alphabet
        \item $K$: a \textbf{finite} set of states
        \item $s\in K$: initial state
        \item $F\subseteq K$: the set of finial states
        \item $\delta$: transition function
        \begin{align*}
            \delta: K \times \Sigma \rightarrow K
        \end{align*}
        $K$ current state, $\Sigma$ symbol, $K$ next state. 
    \end{itemize}
\end{definition}


之前的 symbol 对之后的结果不会有影响.

\begin{definition}[configuration]
    A configuration is any element of $K\times \Sigma^*$
    (current state 与 unread input 作用)
\end{definition}

\begin{definition}[yields in one step]
    $(q,w)\vdash_M (q', w')$ if $w=aw'$ for some $a\in \Sigma$ and $\delta(q, a)=q'$. 
\end{definition}
每走一步, 所携带的输入会减少. 

\begin{definition}[yields]
    $(q,w)\vdash_M^* (q', w')$ if $(q,w)\vdash_M(q', w')$ or $(q,w)\vdash_M\cdots \vdash_M(q', w')$
\end{definition}


\begin{definition}
    $M$ accepts $w\in \Sigma^*$ if $(s,w)\vdash_M^*(q,e)$ for some $q\in F$
\end{definition}

\begin{definition}
    \begin{align*}
        L(M)=\{ w\in \Sigma^*: M\text{ accepts }w \}
    \end{align*}
    $M$ accepts $L(M)$. 
\end{definition}

$M$ accepts $L$ iif
\begin{align*}
    \left\{ \begin{array}{ll}
        w\in L & M\text{ accepts }w\\
        w\notin L & M\text{ doesn't accept }w
    \end{array} \right.
\end{align*}

\begin{definition}
    A language is regular if it's accept by some finite automata. 
\end{definition}

e.g. 
\begin{align*}
    \{ w\in\{ 0,1 \}^*: w \text{ contains $101$ as a substring}\}
\end{align*} is regular? yes.

\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
    \node[state,initial]  (q_0)                      {$q_0$};
    \node[state]          (q_1) [right=of q_0] {$q_1$};
    \node[state]          (q_2) [right=of q_1] {$q_2$};
    \node[state,accepting](q_3) [right=of q_2] {$q_3$};

    \path[->] (q_0) edge node [swap] {1} (q_1)
                    edge [loop above] node {0} ()
              (q_1) edge              node        {0} (q_2)
                    edge [bend right] node [swap] {1} (q_0)
              (q_2) edge node {1} (q_3)
                    edge [bend left] node {0} (q_0)
              (q_3) edge [loop above] node {0,1} ();
\end{tikzpicture}

\subsection{Regular Operations}
\begin{definition}[Regular Operations]\quad

    \begin{itemize}
        \item Union $A\cup B=\{ w: w\in A \text{ or }w \in B \}$
        \item Concatenation $A\cdot B=\{ ab: a\in A\text{ and }b \in B \}$
        \item Star $A^*=\{ w_1 w_2\cdots w_k :w_i \in A\text{ and }k\ge 0 \}$
    \end{itemize}
\end{definition}

regular language 在这些操作下是封闭的. 

\begin{theorem}
    If $A$ and $B$ are regular, so is $A\cup B$. 
\end{theorem}
\begin{proof}
    $\exists M_A=(K_A, \Sigma_A, \delta_A, s_A, F_A)$ accepts $A$, $\exists M_B=(K_B, \Sigma_B, \delta_B, s_B, F_B)$ accepts $B$.
    
    For $M_U=(K_U, \Sigma_U, \delta_U, s_U, F_U)$, 
    \begin{itemize}
        \item $\Sigma_U=\Sigma_A \cup \Sigma_B$
        \item $K_U=K_A\times K_B$
        \item $s_U=(s_A, s_B)$
        \item $F_U=\{ (q_A, q_B) \in K_A\times K_B:q_A\in F_A\text{ or }q_B\in F_B \}$
        \item $\delta_U$: For any $q_A\in K_A$, $q_B\in K_B$, $a\in \Sigma_U$, $\delta_U((q_A, q_B), a)=(\delta_A(q_A,a), \delta_B(q_B, a))$
    \end{itemize}
\end{proof}

\begin{theorem}
    If $A$ and $B$ are regular, os is $A\cdot B$
\end{theorem}
\begin{proof}
    (猜)
\end{proof}

Non-determinism
