\newpage
\section{Computational Complexity}

decidable:
\begin{itemize}
    \item efficiently solvable (tractable)
    \item inefficiently solvable (intractable)
\end{itemize}

resource: time and space

e.g. :$A=\{ 0^k1^k:k\ge 0 \}$

\begin{definition}
    Let $M$ be a deterministic TM that halts on every input. 
    
    The running time f $M$ is a function $f:N\to N$, where on any input of length $n$, $M$ halts within $f(n)$ steps. 
\end{definition}

\begin{definition}
    We say a language $L$ is in DTIME(t(n)) if $L$ is decided by some standard DTM with running time $O(t(n))$
\end{definition}

If Multiple tapes TM is $t(n)$, single tape TM is at most $O(t^2(n))$. 因为多带每走一步, 单带需要以一次扫描确定多带的每一个位置并修改, 扫一遍的时间是 $t(n)$. 

\begin{thesis}[The Cobham-Edmond Thesis]
    Any ``reasonable'' and ``general'' deterministic model of computational is polynomially related. 
\end{thesis}


\begin{definition}
    Let $P$(polynomially decidable) be the set of language that are decided by some DTM with $poly(n)$ time. (robust)
\end{definition}

\begin{theorem}
    Every context-free language is in $P$
\end{theorem}

\begin{proof}
    If $A$ is context-free, $\exists$ CFG $G=(V,\Sigma, S, R)$ in CNF generates $A$, given a string of length $n$, enumerate all derivations of length $\ge n-1$ (total $|R|^{n-1}$)

    Dynamic Programing, 
    \begin{itemize}
        \item $w=a_1 \dots a_n$, $S\Rightarrow^* w$?
        \item subproblem: for $1\le i\le j\le n$, define
        \begin{align*}
            T[i,j]=\{ A\in V-\Sigma : A\Rightarrow^* a_i\dots a_j \}
        \end{align*}
        \item Goal: $T[1,n]$, $S\in T[1,n]$
        \item Base case: for $1\le i\le n$,
        \begin{align*}
            T[i,i]=\{ A\in V-\Sigma: A\to a_i \}
        \end{align*}
        \item Recurrence: for $1\le i\le j\le n$
        \begin{align*}
            T[i,j]=\bigcup_{k=i}^{j-1} \{ A\to BC : B\in T[i,k] \land C\in T[k+1, j] \}
        \end{align*}
        \item $n^2$ subproblems, each subproblem costs: $n\cdot|R|^3$.
    \end{itemize}
    total: $\left( n^3 |R|^3 \right)$, and $R$ is constant
\end{proof}

\subsection{SAT(satisfiability) Problem}

Let $X=\{ x_1,\dots,x_n \}$ be a set of boolean variable. 
\begin{itemize}
    \item $x_1,\dots,x_n,\bar{x}_1,\dots,\bar{x}_n$: literals
    \item a clause is like $x_1\lor x_2 \lor \bar{x}_3$
    \item a boolean formular is like 
    \begin{align*}
        (x_1\lor x_2 \lor \bar{x}_3)\land (x_2\lor x_3 \lor \bar{x}_4)\land (x_1\lor x_3 \lor x_5)
    \end{align*}
    \item A truth assignment of $X$ is $T:X\to \{ 0,1 \}$
\end{itemize}

SAT: given a formular $F$, is there any assignment $T$ that makes $F$ true/ satisfies $F$.

We don't know whether $SAT\in P$ or not.

\subsection{NP}

\begin{definition}
    Let $M$ be a non-deterministic TM such that for any input every branch of $M$ halts within $k$ steps, where $k$ depends only on the input.
    
    The running time of $M$ is a function $f:N\to N$,  for any input of length $n$, every branch of $M$ halts within $f(n)$ steps. 
\end{definition}

SAT can be decided by NTM poly time. 

$M=$ on input $F$
\begin{algorithm}[H]
    \caption{$M$}
    \begin{algorithmic}
        \Require $F$
        \State non-deterministically generate an assignment $T$
        \If{$T$ satisfies $F$}
            \State accept
        \Else
            \State reject
        \EndIf
    \end{algorithmic}
\end{algorithm}

\begin{definition}
    NP is the set of all languages that can be decided by some NTM in poly time. 
\end{definition}

\begin{definition}
    We say that a language A is polynomially verifiable, if there is a polynomially DTM $V$ such that for $x\in\Sigma^*$,
    \begin{enumerate}
        \item if $x\in A$, $\exists\ y$ with $|y|\le poly(|x|)$ s.t. $V$ accepts $\encode{x}\encode{y}$ (completeness)
        \item if $x\notin A$, $\forall\ y$ with $|y|\le poly(|x|)$, $V$ rejects $\encode{x}\encode{y}$ (soundness)
    \end{enumerate}
    called $y$ certificate
\end{definition}

e.g. 
\begin{itemize}
    \item $A=$ SAT
    \item $x=$ boolean formular
    \item $y=$ an truth assignment that satisfies $x$
    \item $V=$ on input $x$ and $y$
    \begin{algorithm}[H]
        \caption{$V$}
        \begin{algorithmic}
            \State evaluate $x$ using $y$
            \If{$x$ is satisfied by $y$}
                \State accept
            \Else
                \State reject
            \EndIf
        \end{algorithmic}
    \end{algorithm}
\end{itemize}

\begin{theorem}
    A language is polynomially verifiable iff it's NP
\end{theorem}
\begin{proof}
    $\Rightarrow$ $\exists$ polynomial-time verifier $V$
    \begin{algorithm}[H]
        \caption{$M$ (If $x\in A$ ?)}
        \begin{algorithmic}
            \Require $x$
            \State non-deterministically generate a certificate $y$ with $|y|\le poly(|x|)$
            \State run $V$ on $\encode{x}\encode{y}$
            \If{$V$ accepts}
                \State accept
            \Else
                \State reject
            \EndIf
        \end{algorithmic}
    \end{algorithm}
    
    $\Leftarrow$ $\exists$ NTM $M$ decides $A$ in $poly(|x|)$ time. 
    
    construct $V$, for $x\in A$, certificate $y=$ branch that accepts $x$, $|y|\le poly(|x|)$ 
    \begin{algorithm}[H]
        \caption{$V$}
        \begin{algorithmic}
            \Require x and y
            \State run $M$ on $x$ under the guidance of $y$
            \If{$M$ accepts}
                \State accept
            \Else
                \State reject
            \EndIf
        \end{algorithmic}
    \end{algorithm}
\end{proof}

$P\subseteq NP$

$P= NP$? i.e. 解决问题的难度与验证问题的难度是否相等. 

widely believe $P\ne NP$

\subsection{NP-complete Problem}
\begin{definition}
    A NP-complete problem is in $p$ iff $P=NP$. 
\end{definition}