\section[The Fundamentals: Algorithms]{The Fundamentals: Algorithms (算法基础)}
\subsection{Algorithms}

\subsubsection{Introduction}

\begin{definition}
    An \highlight{algorithm}: a finite set of precise instructions for performing a computation or solving problem. 
\end{definition}

Pseudocode (伪代码)

\subsubsection{Searching Algorithms}
\begin{enumerate}
    \item Linear search. 
    \item Binary search. 
\end{enumerate}

\subsection{Complexity of Algorithms}
\begin{definition}
    Complexity, space Complexity and time Complexity. 
    \begin{itemize}
        \item \highlight{Complexity}: the amount of time and/or space needed to execute the algorithm. 
        \item \highlight{Space Complexity}: be tied with particular data structures of used to implement the algorithm. 
        \item \highlight{Time Complexity}: can be expressed in terms of the number of operation used by the algorithm when the input has a particular size. 
    \end{itemize}
\end{definition}

Types of Complexity: 
\begin{itemize}
    \item Best-case time
    \item Worst-casr time
    \item Average-case time 
\end{itemize}

\subsubsection{Some Terminology to Describe the Time Complexity}

\begin{center}
    \tablefirsthead{\hline
        Complexity  & Terminology\\ \hline
    }
    \tablehead{\hline
        Complexity  & Terminology\\ \hline
    }
    \tabletail{\hline}
    \tablelasttail{\hline}
    % \tablecaption{}
    \bottomcaption{Commonly used Terminology for the Complexity of Algorithms}
    \begin{supertabular}{|l|r|}
        $O(1)$      &Constant complexity\\
        $O(\log n )$&Logarithmic complexity\\
        $O(n)$      &Linear complexity\\
        $O(n\log n)$&$n \log n$ complexity\\
        $O(n^b)$    &Polynomial complexity\\
        $O(b^n),b>1$&Exponential complexity\\
        $O(n!)$     &Factorial complexity\\ 
    \end{supertabular}
\end{center}

\subsubsection{NP Problem}
\begin{itemize}
    \item P class: Problems can be solved bu Polynomial time algorithm. 
    \item NP class: Problems for which a solution can be checked in Polynomial time. 
    \item NP-Complete Problem: If any of these probelms can be solved bu polynomial worst-case time algorithm, then all can be solvev by polynomial worst-case time algorithms.
\end{itemize}

\begin{figure}[H]
    \centering
    \begin{tikzpicture}
        \node [ellipse, draw=black, text centered] at (-1,0.5) {P Problems};
        \node [ellipse, draw=black, text centered] at (0.8,-0.5) {NP Complete};
        \node at (0,1.2) {NP Problems};
        \draw (0,0.2) ellipse (3.2 and 1.6);

        \draw (-4,-1.6) rectangle (4,2);
    \end{tikzpicture}
    \caption{The World}
\end{figure}
