\section[Induction and Recursion]{Induction and Recursion (归纳与递归)}

\subsection{Mathematical Induction}
\subsubsection{The Well-ordeing Property (良序集)}

Every nonnegative integers has a least element. 

\subsubsection{The Proof by Mathematical Induction}
A proof by mathematical induction that $P(n)$ is true for every positive integer $n$ consists of two steps: 
\begin{enumerate}
    \item Basic step: Th proposition $P(1)$ is shown to be true. 
    \item Inductive step: The implication \\$P(n) \longrightarrow P(n+1)$ is shown to be true for every positive integer $n$. 
\end{enumerate}

\subsubsection{The Second Principle of Mathematical Induction}
\begin{enumerate}
    \item Basic step: Th proposition $P(1)$ is shown to be true. 
    \item Inductive step: The implication \\ $\left[P(1)\land P(2) \land\cdots\land P(n)\right] \longrightarrow P(n+1)$ is shown to be true for every positive integer $n$. 
\end{enumerate}

The two forms of mathematical induction are equivalence. 

\subsection{Recursive Definitions}

\subsubsection{Recursively Defined Functions}
To define a function with the set of nonnegative integers as its domain. 
\begin{enumerate}
    \item Specify the value of the function at zero, $f(0)$.
    \item Give a rule for finding its value as an integer from ist values at smaller integers. 
\end{enumerate}

e.g. The Fibonacci numbers\begin{align*}
    &f_0=1, f_1=1\\
    &f_n=f_{n-1}+f_{n-2}
\end{align*}

\subsubsection{Recursively Defined Sets}
e.g. Let S be defined recursively by 
\begin{enumerate}
    \item $3 \in S$.
    \item $x+y\in S \text{ if } x\in S \text{ and }y \in S$.
\end{enumerate}

$S$ is the set of positive integers divisible by 3. 

The most commom uses is to define well-formed formulae. 

Also can define strings.

\subsubsection{Recursive Algorithms}
An algorithm is called \highlight{recursive} if it solves a problem by reducing it to an instance of the same problem with smaller input. 